---
title: "Homework_10"
author: "Megan O'Connor"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1
```{r}
  counter <- 0
vector <- c(0,1,2,0,7,8,0,9,9,4,5,0)
counting_zeroes <- function(vector) {
  for (i in vector) {
    if (i == 0) {
      counter <- counter + 1
    }
  }
  
  return(counter)
}

counting_zeroes(vector)

```

Question 2
```{r}
zero_subset <- function(vector) {
  return(sum(vector == 0))
}

zero_subset(vector)
```


Question 3: Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}


mtrx <- function(rows,cols) {
  result <- matrix(nrow=rows,ncol=cols)
  
  for (i in 1:rows) {
    for (j in 1:cols) {
      result[i,j] <- i*j
    }
  }
return(result)
}

rows <- 4
cols <- 5
final_matrix <- mtrx(rows,cols)
print(final_matrix)
```

Question 4a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.
```{r}
set.seed(200)
n <- 100
groupmean <- c(10,40,50)

group <- rep(1:3, each = n)
response <- c(rnorm(n, mean = groupmean[1]),
             rnorm(n, mean = groupmean[2]),
             rnorm(n, mean = groupmean[3]))
data <- data.frame(group, response)
print(data)
```

Question 4b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.
```{r}
reshuffle_mean_calc <- function(data) {
  reshuffled_means <- numeric(3)
  for (i in 1:3) {
    group_data <- data[data$group == i, ]
    shuffleresponse <- sample(group_data$response)
    data[data$group == i, "response"] <- shuffleresponse
    reshuffled_means[i] <- mean(shuffleresponse)
    reshuffled_means[i] <- mean(shuffleresponse)
  }
  return(reshuffled_means)
}


calculations <- reshuffle_mean_calc(data)
reshuffle_mean_calc(data)


```

Question 4c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.
```{r}
rerun <- function(data){
  
# Making dataframe
results <- data.frame(matrix(nrow = n, ncol = 4))  # Create an empty matrix to store results
colnames(results) <- c("replicate", paste0("mean_", 1:3))  # Column names for the results


for (i in 1:n) {
  new_means <- reshuffle_mean_calc(data)
  results[i, 1:4] <- c(i,new_means)
  }
(results)
}

n <- 100
rerun(data)


##### OLD UNUSED CODE GRAVEYARD##################
#   # Call the function for each replicate
#   new_data <- data
#   new_means <- reshuffle_mean_calc(new_data)
#   
#   # Store the results in the matrix
#   results[replicate_num, ] <- c(replicate_num, new_means)
# }
# 
# # Convert the matrix to a data frame
# histodata <- as.data.frame(results)
# 
# # Print the first few rows of the results data frame
# print(head(histodata))
```

Question 4d. Use qplot() to create a histogram of the means for each reshuffled group. How do the distributions of reshuffled means compare to the original means?
```{r}
library(ggplot2)
# Create histogram using qplot


qplot(data = calculations, x = reshuffled_means, geom = "histogram",
      bins = 3, fill = I("blue"),
      main = "Histogram of Reshuffled Means",
      xlab = "Mean Value", ylab = "Frequency") +
  geom_vline(xintercept = mean(data$response), color = "red", linetype = "dashed") +
  labs(title = "Comparison of Reshuffled Means to Original Mean",
       subtitle = "Red dashed line indicates the original mean") +
  theme_minimal()
```
