---
title: "Homework_6"
author: "Megan O'Connor"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 6: Simulating and Fitting Data Distributions

### Fake Class Data
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

# quick and dirty, a truncated normal distribution to work on the solution set

# -------------- Read in Data Vector --------------
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

# ----------------- Plot Histogram of Data ------------
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="coral3",size=0.2) 
print(p1)

# -------------- Add empirical density curve ----------
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

# ------------ Max Likelihood Parameters for Normal -----
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

# ----------- NORMAL Prob Density ------------------
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat

# -----------EXPONENTIAL Prob Density ---------------
# models continuous variables
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2

# --------- UNIFORM Prob Density ------------------
# assumes every outcome has same probability of occuring
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3

# ----------- GAMMA Prob Density ---------------
# Sum of expo distrubutions
# time series modeling
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4

# ----------- BETA Prob Density ---------------------
# generalization of uniform distribution
# bounded and continuous, no specific shape
# shape changes based on parameters
# used in bassian analysis, as a way to have an assumption
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="skyblue2",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

### Dolphin data from DRYAD (Allen et. al 2022)
#### "Dynamic body acceleration as a proxy to predict the cost of locomotion in bottlenose dolphins"
```{r}
# This study looked at the energetic cost of locomotion of bottlenose dolphins
# variable I chose to focus on was "ml of O2 per kg dolphin per minute"
z <- read.csv("Fig_3.csv",header=TRUE,sep=",")
str(z)

z<-na.omit(z)
z$myVar <-z$O2_ml_per_kg_min
summary(z)

# Histogram of Data
p1 <- ggplot(data=z, aes(x=myVar, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="royalblue3",size=0.2) 
print(p1)

# Gamma Probability Density (Fits Pretty Well)
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat <- stat_function(aes(x = myVar, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat 

# Normal Probability Density (DOES NOT MATCH WELL!)
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat2 <- stat_function(aes(x = myVar, y = ..y..), fun = dnorm, colour="orchid", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat + stat2
```

### Simulation of Dolphin Data
```{r}
# ------------------ Creating a New Simulation --------------------------
# Max Likelihood Parameters
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] 

# Reading in New Vector
simul <- rnorm(n=3000,mean=5.676607)
simul <- data.frame(1:3000,simul)
names(simul) <- list("ID","myVar")
simul <- simul[simul$myVar>0,]
str(simul)
summary(simul$myVar)

# Histogram of Simulation
p2 <- ggplot(data=simul, aes(x=myVar, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="darkslategrey",size=0.2) 
print(p2)

# Gamma Probability Density of Simulation
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat <- stat_function(aes(x = myVar, y = ..y..), fun = dgamma, colour="brown", n = length(simul$myVar), args = list(shape=shapeML, rate=rateML))
p2 + stat
```

### Final Products!
```{r}
# Fresh Plot of the Original Dolphin Data
print(p1 + stat)

#Fresh Simulated Plot 
print(p2 + stat)
```

#### How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?
Both histograms are similar in scale (by design) however the simulation is much closer to a what a normal distribution curve on the graph, especially because the mean of the original data is close to 5. However, both of their gamma probability density lines had similar shapes, what makes me believe that the simulation still did a good job of simulating realistic data, though the actual data was a bit more extreme, with more outliers than the simulation.

