---
title: "Homework_10"
author: "Megan O'Connor"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1
```{r}
  counter <- 0
vector <- c(0,1,2,0,7,8,0,9,9,4,5,0)
counting_zeroes <- function(vector) {
  for (i in vector) {
    if (i == 0) {
      counter <- counter + 1
    }
  }
  
  return(counter)
}

counting_zeroes(vector)

```

Question 2
```{r}
zero_subset <- function(vector) {
  return(sum(vector == 0))
}

zero_subset(vector)
```


Question 3: Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}


mtrx <- function(rows,cols) {
  result <- matrix(nrow=rows,ncol=cols)
  
  for (i in 1:rows) {
    for (j in 1:cols) {
      result[i,j] <- i*j
    }
  }
return(result)
}

rows <- 4
cols <- 5
final_matrix <- mtrx(rows,cols)
print(final_matrix)
```

Question 4a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.
```{r}
set.seed(200)
n <- 100
groupmean <- c(10,40,50)

group <- rep(1:3, each = n)
response <- c(rnorm(n, mean = groupmean[1]),
             rnorm(n, mean = groupmean[2]),
             rnorm(n, mean = groupmean[3]))
data <- data.frame(group, response)
print(data)
```

Question 4b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.
```{r}
reshuffle <- function(data) {
  for (i in data) {
    shuffleresponse <- sample(response)
  }
  data <- data.frame(group,shuffleresponse)
  return(data)
}


reshuffle(data)
```

Question 4c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.
```{r}
# data frame for my new results
newresults <- data.frame(matrix(nrow=100, ncol=4))
colnames(newresults) <- c("rep_no", "G1_mean", "G2_mean", "G3_mean")

Question4 <- function(data) {
  for (i in 1:nrow(data)){
    newresponse <- reshuffle(data)$shuffleresponse
    newresults[i, 1] <- i 
    newresults[i, 2:4] <- newresponse
  }
  return(newresults)
  }
  
Question4(data)
```

Question 4d. Use qplot() to create a histogram of the means for each reshuffled group. How do the distributions of reshuffled means compare to the original means?
```{r}
library(ggplot2)

# Calculate means of G1, G2, G3


# Create histogram using qplot
qplot(data=newresults, x= , geom="histogram", bins = 3,
      main= "Histogram Reshuffled Means", xlab = "Mean Value", ylab = "Frequency")


```
