---
title: "Homework_4"
author: "Megan O'Connor"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 4: Atomic Vector Puzzles

### Problem 1
First assigned all the variables to their letters. Then assigned z to each formula I needed to make, made sure I used markers like asterisks to create the math problem.
```{r}
x <- 1.1
a <- 2.2
b <- 3.3

z <- x^(a^b)
print(z)
z <- (x^a)^b
print(z)
z <- 3*(x^3)+2*(x^2)+1
print(z)
```

### Problem 2
a.
Combined 2 vectors going forward and backwards
```{r}
z <- c(1,2,3,4,5,6,7,8)
z <- c((seq(from=1, to=8)),(seq(from=7, to=1)))
print(z)
```
b.
Made sure R would only print 1-5, and wanted the numbers to repeat their number of times so made "times" the vector
```{r}
z <- seq(from=1, to=5)
z <- rep(x=z, times=z)
print(z)
```
c.
Similar to b however we wanted the numbers to repeat number of times "backwards" so used "rev" command in the "times" command
```{r}
z <- seq(from=5, to=1)
z <- rep(x=z, times=rev(z))
print(z)
```
### Problem 3
First created the "runifs" to make random numbers for x and y. Then used polar coordinates conversion formula to define r and theta. Then combined those two things to make an actual set of coordinates.
```{r}
x <- runif(1)
print(x)
y <- runif(1)
print(y)

r <- sqrt((x^2)+(y^2))
print(r)
theta <- atan(y/x)
print(theta)

coords <-c(r,theta)
print(coords)
```

### Problem 4
```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)

# Serpent gets in line
queue <- c(queue, "serpent")
print(queue)

# Sheep enters ark
queue <- c(queue[2:5])
print(queue)

# Donkey talks his way to the front
queue <- c("donkey", queue[1:4])
print(queue)

# Serpent gets impatient and leaves
queue <- c(queue[1:4])
print(queue)

# Owl gets bored and leaves
queue <- c(queue[1:2], queue[4])
print(queue)

# Aphid cuts the ant
queue <- c(queue[1:2], "aphid", queue[3])
print(queue)

# Define position of aphid in line
print(which("aphid" == queue))
```

### Problem 5
```{r}
x <- 1:100

y <- which((a%%3!=0) & (a%%3!=0) & (a%%7!=0))
print(y)
```