---
title: "HW_10.RMD"
author: "Megan O'Connor"
date: "2024-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Question 1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}
counting_function <- function(x=my_vec) {
  counter <- 0
  for (i in seq_along(my_vec)) {
    if(my_vec[i]==0) {
      counter <- counter + 1
    }
  }
 return(counter)
  } # end of counting
my_vec <- c(0,1,5,7,0,8,0,5,3,4,1)
counting_function(my_vec)
```
Question 2. Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}
zero_subset <- function(vector) {
  return(sum(vector == 0))
}

zero_subset(vector)
```
Question 3. 
```{r}
mtrx <- function(rows,cols) {
  result <- matrix(nrow=rows,ncol=cols)
  
  for (i in 1:rows) {
    for (j in 1:cols) {
      result[i,j] <- i*j
    }
  }
return(result)
}

rows <- 4
cols <- 5
final_matrix <- mtrx(rows,cols)
print(final_matrix)
```


Question 4a.
```{r}

```
Question 4b. 

Question 4c.

Question 4d.

