---
title: "HW_10.RMD"
author: "Megan O'Connor"
date: "2024-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Question 1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}
  counter <- 0
vector <- c(0,1,2,0,7,8,0,9,9,4,5,0)
counting_zeroes <- function(vector) {
  for (i in vector) {
    if (i == 0) {
      counter <- counter + 1
    }
  }
  
  return(counter)
}

counting_zeroes(vector)
```
Question 2. Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}
zero_subset <- function(vector) {
  return(sum(vector == 0))
}

zero_subset(vector)
```
Question 3. 
```{r}
mtrx <- function(rows,cols) {
  result <- matrix(nrow=rows,ncol=cols)
  
  for (i in 1:rows) {
    for (j in 1:cols) {
      result[i,j] <- i*j
    }
  }
return(result)
}

rows <- 4
cols <- 5
final_matrix <- mtrx(rows,cols)
print(final_matrix)
```


Question 4a.
```{r}
n <- 100
means <- c(50,75,100)
group <- rep(1:3, each=n)
response <- c(rnorm(n,mean=means[1]),
              rnorm(n,mean=means[2]),
              rnorm(n,mean=means[3]))

# making a data fram!
MyData <- data.frame(group, response)
data

```
Question 4b. 
```{r}
# first I need to reshuffle the data and THEN get the means again
reshuffle_time <- function(MyData){
  for (i in MyData) {
    shuffleresponse <- sample(MyData[,2])
    shuffleresponse
    
    # then I make a new frame
    data_2 <- data.frame(group,shuffleresponse)
    data_2
    
    # NOW i get the means
    group_1mean <- mean(data_2[1:n,2])
    group_2mean <- mean(data_2[n:(2*n),2])
    group_3mean <- mean(data_2[(2*n):(3*n),2])
  }
  new_means <-  data.frame(group_1mean, group_2mean, group_3mean)
  return(new_means)
}

reshuffle_time(MyData)
# YAY MY NEW MEANS ARE REAL
```
Question 4c.
```{r}
library(tidyverse)
# using a for loop to replicate these 100 times :)

loopy <- function(MyData) {
  MyResults <- data.frame(matrix(nrow=n,ncol=4))
  colnames(MyResults) <- c("replicate", "group_1mean", "group_2mean", "group_3mean")
  # n = 100 in this case
  for (i in 1:n) {
    new_means <- reshuffle_time(MyData) # bringing back this function
    MyResults[i,1:4] <- c(i,new_means)
  }
  # what are we returning???? RESULTS
  return(MyResults)
}

n = 100 # like i said before

answer <- loopy(MyData)
answer
```
Question 4d.
```{r}
# time to make a nice looking plot that you can fill with whatever group you want to look at
library(ggplot2)
answer

answer_plot <- qplot(data=answer,
              y=answer[,2],
              geom="histogram",
              fill=I("purple"),
              main="Group Means",
              xlim=c(0,100), # repeats 100x
              ylim=c(0,100), # my means go up to 100
              xlab="Replicate",
              # now we let ppl know which group we're looking at
              ylab="Group 1 Mean Value")
answer_plot


```